---
title: "heart failure death predection"
auther: "Waleed Aljomah"
output: html_document
---

# Prereqs
```{r}
library(tidyverse)
```


# [source of data](https://www.kaggle.com/andrewmvd/heart-failure-clinical-data)
# [publication](https://bmcmedinformdecismak.biomedcentral.com/articles/10.1186/s12911-020-1023-5)

Aim of the data according to the publication paper is to predict death in heart failure patients based on 12 features.

features : 

# Sex - Gender of patient Male = 1, Female =0
# Age - Age of patient
# Diabetes - 0 = No, 1 = Yes
# Anaemia - 0 = No, 1 = Yes
# High_blood_pressure - 0 = No, 1 = Yes
# Smoking - 0 = No, 1 = Yes
# DEATH_EVENT - 0 = No, 1 = Yes
# creatinine_phosphokinase
# ejection_fraction
# platelets
# serum_creatinine
# serum_sodium

```{r }
heart <- read.csv("data/heart_failure_clinical_records_dataset.csv")
```

# Explore data


```{r }
head(heart)
summary(heart)
```

# number of patients (death_Event),  death or alive?

```{r}
heart %>%
  mutate(DEATH_EVENT = recode(DEATH_EVENT,
                              `0` = "Alive",
                              `1` = "Death")) %>%
  count(DEATH_EVENT)
```


# plot patients death or alive?


```{r}
heart %>%
  mutate(DEATH_EVENT = recode(DEATH_EVENT,
                              `0` = "Alive",
                              `1` = "Death")) %>%
  ggplot(aes(x = DEATH_EVENT)) +
  geom_bar()
```

# relationship between Death AND Low Ejection fraction AND CPK

#mean of Ejection fraction in all patients


```{r}
mean_Ejection_Fraction_all <- mean(heart$ejection_fraction)
```

#low Ejection fraction in deaths patients only


```{r}
low_Ejection_Fraction_death <- heart %>%
  filter(heart$DEATH_EVENT == '1', heart$ejection_fraction < 40)
```

#mean of Ejection fraction in death and low Ejection fraction
```{r}
mean_Ejection_Fraction_death_Low_EF <-mean(low_Ejection_Fraction_death$ejection_fraction)
```


# plot the relation between death and low ejection fraction? 

```{r}
ggplot(heart, aes(ejection_fraction, fill = factor(DEATH_EVENT))) +
  geom_histogram(aes(y = ..density..), position = position_dodge(width = 0), alpha = 0.5) +
  geom_density(alpha = 0.4, color = NA)
```

# how many patients died with NORMAL ejection fraction AND high CPK > 500 ?


```{r}
heart %>%
  filter(ejection_fraction > 40) %>%
  filter(DEATH_EVENT == '1') %>%
  filter(creatinine_phosphokinase > 500) %>%
  nrow()
```

# how many patients died with NORMAL ejection_fraction and Normal CPK ? 


```{r}
 heart %>%
  filter(DEATH_EVENT == '1') %>%
  filter(ejection_fraction > 40) %>%
  filter(creatinine_phosphokinase < 500) %>%
  nrow()
```

# what is the relation between death and age of the patients? 

# plot Age distribution in all patients 
```{r}
heart %>%
  ggplot(aes(x = age)) +
  geom_bar()
```



#plot Age distribution in deaths only 
```{r}
heart %>%
  filter(DEATH_EVENT == '1') %>%
  ggplot(aes(x = age)) +
  geom_bar()
```


  